---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# filematch

<!-- badges: start -->
<!-- badges: end -->

filematch has tools that aid in constrained statistical matching of two microdata files that have record weights.

These data files commonly are called file A and file B. Typically they have a set of X variables in common. In addition, A has a set of Y variables and B has a set of Z variables. The goal is to create an AB file that has X, Y, and Z variables. 

An unconstrained match does not necessarily ensure that... A constrained match ensures that ...

An excellent package that has many tools for unconstrained file matching is `StatMatch`. 

This package has a different focus that StatMatch: its goal is to ensure that the weighted file AB has weighted sums for the Y variables that match the weighted sums of Y in A, and also has weighted sums of Z variables that match



## Installation

You can install the development version of filematch from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("donboyd5/filematch")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(filematch)
## basic example code
```


